Viscous Erosion in Porous Media

### 2018-02-28 docs/methods/erosion_methods.tex
NM

As a final validation, we now aim to verify that our method converges with second-order accuracy in time. To this end, we run the same single-body simulation from Figs.~\ref{01bodseq} and \ref{shrink_intface} (except with a coarser $N_\iin=256$) to a stopping time of $t_s = 1E-2$. This stopping time is 56\% of the time it takes the body to vanish, and was chosen so that the body will have eroded substantially but is not so small that accuracy is lost. We assess error by measuring the difference between simulations of successive resolutions (i.e.~we verify Cauchy convergence). In particular, we measure the $L^2$-difference of the body shapes obtained at the stopping time. At each stage we halve the timestep $\Delta t$. Table \ref{convtab} shows the $L^2$-errors between successive simulations, along with the order of accuracy as obtained by Richardson-extrapolation-type calculations. The table confirms that our method to converges with second-order accuracy. In the table, we also show the CPU time of the simulation run on a single processor (though the algorithm is easily parallelized), showing a linear increase in computational time.

@@

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{Convergence test for multiple-body erosion with fixed-pressure drop condition. In this test, we use 20 bodies with $N_{\iin} = 128$ grid points on each. We run the simulations to a stopping time of $t_s = 4E-3$, which is 24\% of the time it takes for {\em all} bodies to vanish. We measure the {\em combined} $L^2$-difference in the shapes of {\em all} bodies. The table confirms that the method maintains second-order accuracy in time, even with multiple bodies and a time-varying inflow rate $\umax$.
}
\vspace{0.3 pc}
\label{convtab}
\begin{tabular}{l l l l}
\hline
\hspace{0.0pc} $\Delta t/t_s$
\hspace{0.5pc} & error 
\hspace{0.5pc} & order
\hspace{0.5pc} & CPU time \\
\hline
%
1/10     	& 2.67E-5  	& --        	& 14 mins  	\\
1/20     	& 7.23E-6  	& 1.88 	& 27 mins  	\\
1/40     	& 1.95E-6  	& 1.89 	& 52 mins  	\\
1/80     	& 4.90E-7  	& 1.99 	& 1.7 hours	\\
1/160     	& 1.25E-7  	& 1.97  	& 3.4 hours	\\
%
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Parameters in this test are: 20circ128, nits = 5, dt0 = 4e-4, tfin = 4e-3
% epsfac = 10, sigfac = 10, fixarea = 0, fixprdop = 1

@@

% Table: GMRES iterations
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{GMRES Iterations
}
\vspace{0.3 pc}
\label{itertab}
\begin{tabular}{c c | c c}
\hline
\hspace{0.5pc} Number of bodies
\hspace{0.5pc} & GMRES iterations 
\hspace{0.5pc} & Resolution, $N_{\iin}$
\hspace{0.5pc} & GMRES iterations  \\
\hspace{0.0pc} ($N_\iin = 128$ in each) & 
\hspace{0.5pc} &(10 bodies in each) & \\
\hline
%
10	& 223	& 32	    	& ?	\\
15	& 292	& 64    	& ?	\\
20	& 363	& 128	& ?	\\
30	& 496	& 256	& ?	\\
40	& 676	& 512	& ?	\\
50	& 991	& 1024	& ?	\\
%
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% In these tests I always use 128 points per body.

>>>>

As a final validation, we now aim to verify that our method converges with second-order accuracy in time. To this end, we run the same single-body simulation from Figs.~\ref{01bodseq} and \ref{shrink_intface} (except with a coarser $N_\iin=256$) to a stopping time of $t_s = 1E-2$. This stopping time is 56\% of the time it takes the body to vanish, and was chosen so that the body will have eroded substantially but is not so small that accuracy is compromised. We assess error by measuring the difference between simulations of successive resolutions (i.e.~we verify Cauchy convergence). In particular, we measure the $L^2$-difference of the body shapes obtained at the stopping time. At each stage we halve the timestep $\Delta t$. Table \ref{convtab} shows the $L^2$-errors between successive simulations, along with the order of accuracy as obtained by Richardson-extrapolation-type calculations. The table confirms that our method to converges with second-order accuracy. In the table, we also show the CPU time of the simulation run on a single processor (though the algorithm is easily parallelized), showing a linear increase in computational time.

@@

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{Convergence test for multiple-body erosion with fixed-pressure-drop condition. In this test, we use 20 bodies with $N_{\iin} = 128$ grid points on each. We run the simulations to a stopping time of $t_s = 4E-3$ (24\% of the time it takes for {\em all} bodies to vanish). We measure the {\em combined} $L^2$-difference in the shapes of {\em all} bodies. The table confirms that the method maintains second-order accuracy in time, even with multiple bodies and a time-varying inflow rate $\umax$.
%We compare the shapes of {\em all} bodies from simulations of successive resolution using {\combined} $L^2$-norm
}
\vspace{0.3 pc}
\label{convtab}
\begin{tabular}{l l l l}
\hline
\hspace{0.0pc} $\Delta t/t_s$
\hspace{0.5pc} & error 
\hspace{0.5pc} & order
\hspace{0.5pc} & CPU time \\
\hline
%
1/10     	& 2.67E-5  	& --        	& 14 mins  	\\
1/20     	& 7.23E-6  	& 1.88 	& 27 mins  	\\
1/40     	& 1.95E-6  	& 1.89 	& 52 mins  	\\
1/80     	& 4.90E-7  	& 1.99 	& 1.7 hours	\\
1/160     	& 1.25E-7  	& 1.97  	& 3.4 hours	\\
%
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Parameters in this test are: 20circ128, nits = 5, dt0 = 4e-4, tfin = 4e-3
% epsfac = 10, sigfac = 10, fixarea = 0, fixprdop = 1

@@

% Table: GMRES iterations
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{GMRES Iterations
}
\vspace{0.3 pc}
\label{itertab}
\begin{tabular}{c c | c c}
\hline
\hspace{0.5pc} Number of bodies
\hspace{0.5pc} & GMRES iterations 
\hspace{0.5pc} & Resolution, $N_{\iin}$
\hspace{0.5pc} & GMRES iterations  \\
\hspace{0.0pc} ($N_\iin = 128$ fixed) & 
\hspace{0.5pc} &(10 bodies) & \\
\hline
%
10	& 223	& 32	    	& ?	\\
15	& 292	& 64    	& ?	\\
20	& 363	& 128	& ?	\\
30	& 496	& 256	& ?	\\
40	& 676	& 512	& ?	\\
50	& 991	& 1024	& ?	\\
%
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% In these tests I always use 128 points per body.

With the multiple-body simulations validated, we now discuss the emergent physics of viscous erosion with multiple bodies...
